disagapp Session 2024-11-07
================

Please find below the R code history from your *disagapp* v1.0.0
session.

You can reproduce your session results by running this R Markdown file
in RStudio.

Each code block is called a “chunk”, and you can run them either
one-by-one or all at once by choosing an option in the “Run” menu at the
top-right corner of the “Source” pane in RStudio. The file can also be
rendered into an html file using `rmarkdown::render()` which will
contain all of the outputs alongside the code used to generate them.

For more detailed information see <http://rmarkdown.rstudio.com>.

### Package installation

disagapp uses the following R packages that must be installed and loaded
before starting.

```{r}
library(terra)
library(leaflet)
library(disaggregation)
library(disagapp)

#create empty list to store covariate data
covariates <- list()

#set a seed
set.seed(319)
```

The *disagapp* session code .Rmd file is composed of a chain of module
functions that are internal to *disagapp*. Each of these functions
corresponds to a single module that the user ran during the session. To
see the internal code for these module functions, click on the links in
the .Rmd file. Users are encouraged to write custom code in the .Rmd
directly to modify their analysis, and even modify the module function
code to further customize. To see the source code for any module
function, just type its name into the R console and press Return.

Your analyses are below.

------------------------------------------------------------------------

Load the selected example dataset

```{r}
dataset <- "mad"
switch(dataset,
       "mad" = {
         shpdf <- data.frame(datapath = list.files(system.file("extdata", "shapes", package="disagapp"), full.names = TRUE),
                             name = list.files(system.file("extdata", "shapes", package="disagapp")))
         shape <- resp_shape(shpdf)
         },
       "nys" = {
         shape <- SpatialEpi::NYleukemia_sf
       },
       "scot" = {
         shape <- SpatialEpi::scotland_sf
         shape$geometry <- shape$geometry * 1000
         shape <- sf::st_set_crs(shape, 27700)
         shape <- sf::st_transform(shape, crs = 4326)
       }
)
```

Load accessibility data and add to covariates list

```{r}
access <- disagapp::cov_access(shape, "Travel Time to Cities (2015)")
covariates[["Travel Time to Cities (2015)"]] <- access
```

Download bioclim data and add to covariates list

```{r}
bioclim <- disagapp::cov_bioclim(shape, "MDG", c("Mean temperature", "Total precipitation"))
covariates <- append(covariates, bioclim)
```

Download land use data and add to covariates list

```{r}
land_use <- disagapp::cov_landuse(shape, 2019, c("BuiltUp", "Crops"))
covariates <- append(covariates, land_use)
```

Download population count data from Worldpop

```{r}
aggregation <- disagapp::agg_worldpop(shape, "MDG", "Constrained", "1km", 2020)
```

Build the spatial mesh

```{r}
mesh <- disaggregation::build_mesh(shape, 
                                   mesh_args = list(
                                    convex = -0.01,
                                    concave = -0.5,
                                    resolution = 300,
                                    max.edge = c(1.5, 3.1), 
                                    cutoff = 0.05,
                                    offset = c(1.5, 3.1)))
```

Prepare the covariates and aggregation raster so that their extent and
resolution match, enabling them to be stacked into a single SpatRaster

```{r}
#temporarily add the aggregation raster to the covariates list
covariates$aggregation <- aggregation

#prepare the summary
covariate_summary <- prep_summary(covariates, remove = FALSE)

#show the table
DT::datatable(covariate_summary)
```

```{r}
covariates$aggregation <- NULL
covariates_prepared <- lapply(covariates, terra::resample, covariates[["BuiltUp land use"]])

#store and then remove the prepared aggregation raster
aggregation_prepared <- terra::resample(aggregation, covariates[["BuiltUp land use"]], method = "sum")

#convert the list of SpatRasters into a multi-layered SpatRaster
covariates_prepared <- terra::rast(covariates_prepared)
```

Scale the covariates and store the parameters for potential reuse

```{r}
scaled_covariate_output <- prep_scale(covariates_prepared)

covariates_prepared <- scaled_covariate_output[["covariates"]]
scaling_parameters <- scaled_covariate_output[["parameters"]]
```

Reduce resolution of covariates and aggregation raster

```{r}
covariates_prepared <- terra::aggregate(covariates_prepared, fact = 4, fun = "mean")
aggregation_prepared <- terra::aggregate(aggregation_prepared, fact = 4, fun = "sum")
```

Plot a correlation matrix of the prepared covariates

```{r}
correlation_matrix <- disagapp::prep_correlation(covariates_prepared)

corrplot::corrplot(correlation_matrix, 
                   method = "circle",
                   type = "lower",
                   diag = FALSE)
```

Prepare the data for fitting the model and plot a summary.

```{r}
prepared_data <- disaggregation::prepare_data(polygon_shapefile = shape,
                                     covariate_rasters = covariates_prepared,
                                     aggregation_raster = aggregation_prepared,
                                     id_var = "ID_2",
                                     response_var = "inc",
                                     na_action = TRUE,
                                     make_mesh = FALSE)

prepared_data$mesh <- mesh

plot(prepared_data)
```

Fit the model

```{r}
# Create NULL priors if they have not been set
if (!exists("priors")) {
  priors <- NULL
}

fitted_model <- disaggregation::disag_model(data = prepared_data,
                                            priors = priors,
                                            family = "poisson",
                                            link = "log",
                                            iterations = 100,
                                            field = TRUE,
                                            iid = TRUE
                                            )
```

Generate predictions from the model

```{r}
prediction <- disaggregation::predict_model(fitted_model, predict_iid = TRUE)

plot(prediction$prediction)

prediction$cases <- prediction$prediction * aggregation_prepared

plot(prediction$cases)

if (!is.null(prediction$field)){
  terra::crs(prediction$field) <- terra::crs(prepared_data$covariate_rasters[[1]])
  prediction$field <- terra::mask(prediction$field, prepared_data$covariate_rasters[[1]])
  plot(prediction$field)  
}

if (!is.null(prediction$iid)){
  plot(prediction$iid)  
}
```
